---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# babynamesIL

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/babynamesIL)](https://CRAN.R-project.org/package=babynamesIL)
[![R-CMD-check](https://github.com/aviezerl/babynamesIL/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/aviezerl/babynamesIL/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `babynamesIL` package provides a comprehensive dataset of names given to Israeli babies from 1948 to 2023. This data, sourced from the Israeli Central Bureau of Statistics (LAMAS), includes names used for at least 5 children in at least one gender and sector (Jewish, Muslim, Christian, Druze, and Other).

## Available Data

The package offers two main data objects:

- `babynamesIL`: A dataset containing the number of babies given each name, categorized by year, sex, and sector from 1948 to 2023.
- `babynamesIL_totals`: A summary dataframe with the total number of babies given each name, aggregated by sex and sector.

## Installation

``` r
# Install the released version from CRAN
install.packages("babynamesIL")

# Install the development version from GitHub
devtools::install_github("aviezerl/babynamesIL")
```

## Example

```{r example}
library(babynamesIL)

# yearly data
head(babynamesIL)

# totals per name
head(babynamesIL_totals)
```

Add the missing values: 

```{r example-missing}
library(dplyr)
babynamesIL %>%
    tidyr::complete(sector, year, sex, name, fill = list(n = 0, prop = 0))
```

Plot trajectory of names: 

```{r example-traj}
library(ggplot2)
babynamesIL %>%
    tidyr::complete(sector, year, sex, name, fill = list(n = 0, prop = 0)) %>%
    filter(name == "מעיין", sector == "Jewish") %>%
    ggplot(aes(x = year, y = prop, color = sex)) +
    geom_line() +
    ggsci::scale_color_aaas() +
    scale_y_continuous(labels = scales::percent) +
    ggtitle("מעיין הבן מול מעיין הבת") +
    theme_classic()
```


## Data Sources 

Data was downloaded from: [here](https://www.cbs.gov.il/he/publications/LochutTlushim/2020/%D7%A9%D7%9E%D7%95%D7%AA-%D7%A4%D7%A8%D7%98%D7%99%D7%99%D7%9D.xlsx). 

2022 data was downloaded from: [here](https://www.cbs.gov.il/he/mediarelease/Pages/2023/%D7%94%D7%A9%D7%9E%D7%95%D7%AA-%D7%94%D7%A4%D7%A8%D7%98%D7%99%D7%99%D7%9D-%D7%A9%D7%A0%D7%99%D7%AA%D7%A0%D7%95-%D7%9C%D7%99%D7%9C%D7%99%D7%93%D7%99-2022.aspx)

2023 data was downloaded from: [here](https://www.cbs.gov.il/he/mediarelease/Pages/2024/%D7%94%D7%A9%D7%9E%D7%95%D7%AA-%D7%94%D7%A4%D7%A8%D7%98%D7%99%D7%99%D7%9D-%D7%A9%D7%A0%D7%99%D7%AA%D7%A0%D7%95-%D7%9C%D7%99%D7%9C%D7%99%D7%93%D7%99-2023.aspx)

## Web App 

A web app using the data is available [here](http://babynames.lifshitz.xyz)
