---
title: "analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(babynamesIL)
library(tidyverse)
library(tgstat)
theme_set(theme_classic())
```

# Israeli baby names

## Distribution of names

We will start by looking at the distribution total number of babies for each name: 

```{r, fig.width = 10, fig.height = 5}
babynamesIL_totals %>%
  mutate(sector = factor(sector, levels = c("Jewish", "Muslim", "Christian", "Druze", "Other"))) %>% 
  ggplot(aes(x = total, color = sex)) +
  ggsci::scale_color_aaas() +
  geom_density() +
  scale_x_log10() +
  facet_grid(.~sector) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Clustering 

We will then create a matrix of the names and their frequencies over time for Jewish female babies. 

```{r}
names_mat <- babynamesIL %>%
        filter(sector == "Jewish", sex == "F") %>%
        select(year, name, prop) %>%
        spread(year, prop, fill = 0) %>%
        column_to_rownames("name") %>%
        as.matrix()
dim(names_mat)
```


Normalize each name: 

```{r}
mat_norm <- names_mat / rowSums(names_mat)
```

Select only names with at least 500 babies: 

```{r}
mat_norm_f <- mat_norm[babynamesIL_totals %>% filter(sector == "Jewish", sex == "F") %>% filter(total >= 500) %>% pull(name), ]
dim(mat_norm_f)
```

Cluster: 

```{r}
hc <- tgs_cor(t(mat_norm_f)) %>% tgs_dist() %>% hclust(method = "ward.D2")
```

Reorder the clustering by year: 

```{r}
hc <- as.hclust(reorder(
  as.dendrogram(hc),
  apply(mat_norm_f, 1, which.max),
  agglo.FUN = mean
))
```

Plot the matrix: 

```{r, fig.width = 5, fig.height = 12}
text_mat <- babynamesIL %>%
  filter(sector == "Jewish", sex == "F") %>%
  tidyr::complete(sector, year, sex, name, fill = list(n = NA)) %>%
  mutate(text = paste(name, paste0("year: ", year), paste0("n: ", n), sep = "\n")) %>%
  select(year, name, text) %>%
  spread(year, text) %>%
  column_to_rownames("name") %>%
  as.matrix()
plotly::plot_ly(z = mat_norm_f[hc$order, ], y = rownames(mat_norm_f)[hc$order], x = colnames(mat_norm_f), type = "heatmap", colors = colorRampPalette(c("white", "blue", "red", "yellow"))(1000), hoverinfo = "text", text = text_mat[hc$order, ]) %>%
  plotly::layout(yaxis = list(title = ""), xaxis = list(title = "Year"))
```

We will wrap it all in a function: 

```{r}
cluster_names <- function(sector, sex, min_total = 500, colors = colorRampPalette(c
  ("white", "blue", "red", "yellow"))(1000)){
  names_mat <- babynamesIL %>%
    filter(sector == !!sector, sex == !!sex) %>%
    select(year, name, prop) %>%
    spread(year, prop, fill = 0) %>%
    column_to_rownames("name") %>%
    as.matrix()
  text_mat <- babynamesIL %>%
    filter(sector == !!sector, sex == !!sex) %>%
    tidyr::complete(sector, year, sex, name, fill = list(n = NA)) %>% 
    mutate(text = paste(name, paste0("year: ", year), paste0("n: ", n), sep = "\n")) %>% 
    select(year, name, text) %>%
    spread(year, text) %>%
    column_to_rownames("name") %>%
    as.matrix()
  mat_norm <- names_mat / rowSums(names_mat)
  mat_norm_f <- mat_norm[babynamesIL_totals %>%
    filter(sector == !!sector, sex == !!sex) %>%
    filter(total >= min_total) %>%
    pull(name), ]
  text_mat <- text_mat[rownames(mat_norm_f), colnames(mat_norm_f)]
  hc <- tgs_cor(t(mat_norm_f)) %>%
    tgs_dist() %>%
    hclust(method = "ward.D2")
  hc <- as.hclust(reorder(
    as.dendrogram(hc),
    apply(mat_norm_f, 1, which.max),
    agglo.FUN = mean
  ))
  plotly::plot_ly(z = mat_norm_f[hc$order, ], y = rownames(mat_norm_f)[hc$order], x = colnames(mat_norm_f), type = "heatmap", colors = colors, hoverinfo = "text", text = text_mat[hc$order, ]) %>%
    plotly::layout(yaxis = list(title = ""), xaxis = list(title = "Year"))
}
```

We can now plot also the Male names: 

```{r, fig.width = 5, fig.height = 12}
cluster_names("Jewish", "M")
```

Or other sectors: 

```{r, fig.width = 5, fig.height = 12}
cluster_names("Muslim", "M")
cluster_names("Muslim", "F")
cluster_names("Christian", "M", 50)
cluster_names("Christian", "F", 50)
cluster_names("Druze", "M", 50)
cluster_names("Druze", "F", 50)
cluster_names("Other", "M", 50)
cluster_names("Other", "F", 50)
```